\chapter{Some useful results}
\section{Lower bounds for the order of $K_2(\mathbb{Z}G)$ and $Wh_2(G)$} % (fold)
\label{sec:lower_bounds_for_the_order_of_}
2016.3.19 阅读这篇文章 1976年发表在\em{Math. Ann.}。

基本假设：$p$: rational prime, $G$: elementary abelian $p$-group.

用的方法：Bloch; van der Kallen  $K_2$ of truncated polynomial rings

结论：
the $p$-rank of $K_2(\mathbb{Z}G)$ grows expotentially with the rank of $G$.

$Wh_2(G)$: ``pseudo-isotopy'' group is nontrivial if $G$ has rank at least $2$.


这篇文章之前已知的结论
Dunwoody, $G$ cyclic of order 2 or 3, $K_2(\mathbb{Z}G)$ is an elementary abelian $2$-group of rank 2 if $G$ has order $2$ and of rank $1$ if $G$ has order $3$. 两者都有$Wh_2(G)$平凡。


一些要知道的东西
$R$ commutative ring, $A$ a subring of $R$.
$\Omega_{R/A}^1$ the module of K\"{a}hler differentials of $R$ considerd as an algebra over $A$ and $R^*$ will denote the group of units of $R$.

the $p$-rank of an abelian group $G$ is $\dim_{\mathbb{F}_p}(G\otimes_\mathbb{Z} \mathbb{F}_p)$.

\\paragraph{第一部分} % (fold)
\label{par:第一部分}
环是$\mathbb{F}_q$ 有限域的情况。

先说结论

首先是一个奇素数的结论
\begin{prop}
	Let $q=p^f$ be odd and let $G$ be an elementary abelian $p$-group of rank $n$.
	Then $K_2(\mathbb{F}_qG)$ is an elementary $p$-group of rank $f(n-1)(p^n-1)$.
\end{prop}
接着是素数$2$的结论
\begin{prop}
	Let $q=2^f$ be odd and let $G$ be an elementary abelian $2$-group of rank $n$.
	Then $K_2(\mathbb{F}_qG)$ is an elementary $2$-group of rank $f(n-1)(2^n-1)$.
\end{prop}

结论实际上是可以统一的，但是方法有些区别，因此原文中分开表述。

我们引进方法时借鉴了van der Kallen的方法和记号


\paragraph{第二部分} % (fold)
\label{par:第二部分}
第二部分是考了系数环是$\mathbb{Z}$的情形，如何将上面的有限域和这里的整数环联系起来，就是用了一个相对$K$群的正合列。
% paragraph 第二部分 (end)
% paragraph 第一部分 (end)
















% section lower_bounds_for_the_order_of_ (end)